---
title: Add project title here
subtitle: Add project subtitle here (or remove this field)
execute:
  echo: true
---

Add 1-3 sentence project abstract here.

## Introduction and data

This section includes an introduction to the project motivation, data, and research question.
Describe the data and definitions of key variables.
It should also include some exploratory data analysis.
All of the EDA won’t fit in the paper, so focus on the EDA for the response variable and a few other interesting variables and relationships.

The gender pay gap has always been a relevant issue for women seeking equal workplace rights.
Given the current political state and changes to DEI policies, this topic is crucial to understand and address from a statistical perspective.
Our research aims to analyze how women are marginalized in various job settings and categories, specifically regarding how pay gaps fluctuate based on how male-dominated a particular field is.
We are interested in how pay gaps vary across time from 2013 to 2016 and within specific major and minor job categories, such as the Management, Business, and Financial sector.
We hypothesize that occupations that tend to be “male-dominated” will have a more significant male-female pay gap and fewer women working in that career.
Our research question asks, to what extent does the gender pay gap vary between male-dominated and female-dominated occupations?
We will primarily consider the variables of earnings, occupation type, and gender representation in the workforce from 2013 to 2016.

All data was collected between 2013 and 2016 from statistics provided by the U.S.
Bureau of Labor through the U.S.
Census.
There are 2089 observations total, with 12 columns.
Each observation is a different combination of the year observed and occupation (i.e., “2013” and “Legislators” is one row/observation).
The variables of interest are the year, occupation, observation categories, how many workers there are in that occupation, the proportion of female workers, total earnings, and the percentage of a male wage that a woman earns.
Male-dominated fields will be specified through the creation of two binary variables: male_dom_workforce labels observations with "Yes" when the number of male workers is greater than the number of female workers and "No" when there are more female workers and male_dom_earnings labels observations with "Yes" when the total male earnings are greater than the total female earnings and "No" when there are higher female earnings.
[**We have also created a numerical variable of perc_male_earnings that measures the percentage of total workforce earnings that are from male employees.**]{.underline} We are also creating the variable median_earnings_gap to measure the difference between male and female wages.
If the value is greater than zero, males tend to be paid more in their positions.
We also added variables that simply transform columns to provide more information and make data visualization and manipulation easier.
One of those is percent_male which subtracts the percentage of female employees from 1 to calculate the percentage of male employees.
We also made percent_female and wage_percent_of_male into decimals rather than numbers.
Finally, we renamed the total_earnings columns to median_earn,ings given that the word "total" was misleading in communicating what the columns represented.

Considering that this data was scrubbed of any identifying factors aside from gender and was made public by the U.S.
Bureau of Labor Statistics, there are no ethical concerns with the data itself.

```{r}
#| label: load-data-jobs-gender
library(tidyverse)
jobs_gender <- read_csv("data/jobs_gender.csv")
glimpse(jobs_gender)
```

```{r}
#| label: mutate-variables-to-decimal-form-and-rename
# mutating percent_female to be displayed as a decimal
jobs_gender <- jobs_gender |>
  mutate(
# mutating percent_female to be displayed as a decimal
    percent_female = (percent_female * 0.01),
# mutating wage_percent_of_male to be displayed as a decimal
    wage_percent_of_male = (wage_percent_of_male * 0.01),
  ) |>
  rename(
# renaming total_earnings_male to median_earnings_male to be more clear
    median_earnings_male = total_earnings_male,
# renaming total_earnings_female to median_earnings_female to be more clear
    median_earnings_female = total_earnings_female,
# renaming total_earnings to median_earnings to be more clear
    median_earnings = total_earnings
  )
```

```{r}
#| label: creating-new-variables
jobs_gender <- jobs_gender |>
  mutate(
# creating wage_percent_of_female 
    wage_percent_of_female = (1 - wage_percent_of_male),
# creating median_earnings_gap to measure the difference between male and female wages. If the value is greater than zero, males tend to be paid more in their position.
    median_earnings_gap = (median_earnings_male - median_earnings_female),
# calculate percentage of male workers in each occupation
    percent_male = (1 - percent_female),
# calculate perc_workforce_gap to calculate how much above or below the male employment is from a "middle ground" of 0.5.
    perc_workforce_gap = percent_male - 0.5,
# create male_dom_workforce as a binary factor where "Yes" means that there are more male workers than female workers, and "No" when there are more female workers.
    male_dom_workforce = as.factor(if_else(
      workers_male > workers_female, "Yes", "No")),
# create male_dom_earnings as a binary factor where "Yes" means that there are  male workers tend to be paid more than female workers, and "No" when female workers are usually paid more.
    male_dom_earnings = as.factor(if_else(
      median_earnings_male > median_earnings_female, "Yes", "No"))
    )
```

```{r}
#| label: scatterplot-wage-percentage-male
#| fig-asp: 1

jobs_gender |>
  ggplot(aes(x = year, y = wage_percent_of_male, color = major_category)) +
  geom_jitter()

```

## Methodology

This section includes a brief description of your analysis process.
Explain the reasoning for the types of analyses you do, exploratory, inferential, or modeling.
If you’ve chosen to do inference, make sure to include a justification for why that inferential approach is appropriate.
If you’ve chosen to do modeling, describe the model(s) you’re fitting, predictor variables considered for the model including any interactions.
Additionally, show how you arrived at the final model by describing the model selection process, interactions considered, variable transformations (if needed), assessment of conditions and diagnostics, and any other relevant considerations that were part of the model fitting process.

## Results

This is where you will discuss your overall finding and describe the key results from your analysis.
The goal is not to interpret every single element of an output shown, but instead to address the research questions, using the interpretations to support your conclusions.
Focus on the variables that help you answer the research question and that provide relevant context for the reader.

## Discussion

In this section you’ll include a summary of what you have learned about your research question along with statistical arguments supporting your conclusions.
In addition, discuss the limitations of your analysis and provide suggestions on ways the analysis could be improved.
Any potential issues pertaining to the reliability and validity of your data and appropriateness of the statistical analysis should also be discussed here.
Lastly, this section will include ideas for future work.
