---
title: Differences in the Gender Pay Gap Across Job Categories and Female Workforce Representation
subtitle: 2013-2016 trends in female wages as a percent of male wages
execute:
  echo: false
  message: false
  warning: false
---

Our research aims to analyze how women are marginalized in various job settings, specifically regarding how pay gaps fluctuate based on the type of job and how male-dominated a particular workforce is.

## Introduction and Exploratory Data

The gender pay gap has always been a relevant issue for women seeking equal workplace rights.
As will be mentioned later, our data only spans from 2013-2016, yet gender pay discrepancies are nothing new.
The gender pay gap narrowed a mere 4% from 2003 to 2024, with women earning 19% less than men in 2003 and 15% less in 2024 (Fry & Aragão, 2025).
The closing of this wage gap is slow-moving, and strides towards resolution and full equality do not seem close.
Some studies estimate that it can take anywhere between 63 to 134 years for women to earn as much as men for the same job (Holder 2025, The Gender).
Not only are these discrepancies unfair in the short term, but they also impact working women post-retirement: when paid less, women end up receiving less in Social Security and pensions and up to 30% less than men in retirement funds for the same job and number of years worked (The Simple).
With the government’s recent changes to DEI policies, understanding workplace wage equality (or lack thereof) from a statistical perspective has never been more crucial to making strides in a progressive, educated direction.
Thus, our research aims to explore how gender pay gaps vary by female representation in the workforce and specific major job categories (such as the Management, Business, and Financial sector).

Our research question asks, **how does the gender pay gap vary between male-dominated and female-dominated workforces and across major job categories?**

We will primarily consider the variables of female wages as a percent of male wages (`wage_percent_of_male`), major job category (`major_category`), and percentage of females in the workforce (`percent_female`).
To clarify-- if women make 90 cents for every dollar men make doing the same job, then `wage_percent_of_female` will equal 0.90. This is also known as the wage gap.

All data was collected between 2013 and 2016 from statistics provided by the U.S.
Bureau of Labor through the U.S.
Census.
There are 2089 observations total, with 12 columns.
Each observation is a different combination of the year observed and occupation (i.e., “2013” and “Legislators” is one row/observation, and there are four rows of "Legislators" across all years).
301 occupations are shown throughout the 4 years, and 22 occupations are not represented in every year. We will refer to observations as "workforces" to maintain clarity in this paper.
Considering that this data was scrubbed of any identifying factors aside from gender and was made public by the U.S.
Bureau of Labor Statistics, there are no ethical concerns with the data itself.

We transformed column names to provide more information and make data visualization and manipulation easier.
Those transformations include renaming the `total_earnings`, `total_earnings_male`, and `total_earnings_female` columns to `median_earnings`, `median_earnings_male`, and `median_earnings_female` given that the word "total" was misleading in communicating that the column represented the median earnings rather than all earnings for each group.
We also made `percent_female` and `wage_percent_of_male` into decimals rather than percentages.
Finally, to make the data easier to work with, we omitted observations with N/A values.

After our data cleaning, we have considered the following variables to be important in our data analysis: `major_category` (broad occupation category), `workers_male` and `workers_female` (estimated male or female full time workers above 16 years old), `percent_female` (percent of females for specific occupations), `total_earnings_male` which became `median_earnings_male` and `total_earnings_female` which became `median_earnings_female` (estimated male or female median earnings for full time workers over 16), and `wage_percent_of_male` (female wages as percent of male wages).

```{r}
#| label: load-packages
#| message: FALSE
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(dplyr)
jobs_gender <- read_csv("data/jobs_gender.csv")
```

```{r}
#| label: mutate-variables-to-decimal-form-and-rename
#| message: FALSE
# mutating percent_female to be displayed as a decimal
jobs_gender <- jobs_gender |>
  mutate(
# mutating percent_female to be displayed as a decimal
    percent_female = (percent_female * 0.01),
# mutating wage_percent_of_male to be displayed as a decimal
    wage_percent_of_male = (wage_percent_of_male * 0.01)
  ) |>
  rename(
# renaming total_earnings_male to median_earnings_male to be more clear
    median_earnings_male = total_earnings_male,
# renaming total_earnings_female to median_earnings_female to be more clear
    median_earnings_female = total_earnings_female,
# renaming total_earnings to median_earnings to be more clear
    median_earnings = total_earnings
  )
```

```{r}
#| label: creating-new-variables
#| message: FALSE
jobs_gender <- jobs_gender |>
# omit all observations with NA values.
  na.omit(jobs_gender) |>
  mutate(
# create male_dom_workforce as a binary factor where "Male dominated" means that there are more male workers than female workers, and "Female dominated" when there are more female workers.
    male_dom_workforce = as.factor(if_else(
      workers_male > workers_female, "Majority male", "Majority female")),
# create male_dom_earnings as a binary factor where Yes" means that male workers tend to be paid more than female workers, and "No" when female workers are usually paid more.
    male_dom_earnings = as.factor(if_else(
      median_earnings_male > median_earnings_female, "Yes", "No"))
    )
```

In order to begin exploring the data, we created the following three visualizations to understand the distribution of wage inequality across job categories, the proportion of male-majority workforces in each job category, and the variation in how much women are paid compared with men. 

Based on Figure 1, we observed that all observations where women are paid equally or more than men are outside of the 75th percentile of all job categories. We also learned that computer, engineering, and science professions have the smallest spread in wage inequality.

```{r}
#| label: boxplot-wage-percentage-male
#| fig-asp: 0.5
#| fig-height: 8
jobs_gender |>
  ggplot(aes(x = wage_percent_of_male, color = major_category)) +
  geom_boxplot() +
  guides(color = guide_legend(reverse = TRUE)) +
  labs(
    title = "Figure 1: 
Distribution of female wages as percent of male wages",
    subtitle = "by major job category",
    x = "Female wages as a percent of male wages",
    color = "Major category"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(axis.ticks.y = element_blank(),
  axis.text.y = element_blank())
```

Male-dominated fields will be specified through the creation of two binary variables: `male_dom_workforce` labels observations with "Male majority" when the number of male workers is greater than the number of female workers and "Female majority" when there are more female workers, and `male_dom_earnings` labels observations with "Yes" when the median male earnings are greater than the median female earnings and "No" when there are higher female earnings.
The following table demonstrates that, regardless of year, the amount of workforces showing male-dominated earnings in this study remains around 300, and that occupations with female-dominated earnings, while showing a slight increase over time, are still under 11.
Therefore, we do not put emphasis on year in this study as numbers remain fairly constant.

*Table 1*

```{r}
#| label: eda-male-dom-earnings
#| message: FALSE
eda_male_dom_earnings <- jobs_gender |>
  group_by(male_dom_earnings, year) |>
  count()
kableExtra::kable(eda_male_dom_earnings)
```

Figure 2 shows the proportion of occupations within the different major categories that reported having a male- or female-dominated work force.
The dashed line shows the 50% mark in the proportions, so any bar showing the blue male-majority line surpassing the dashed line has majority male-dominated work forces.
The graph shows that there is the least amount of female representation across observations in the major job category of "Natural Resources, Construction, and Maintenance", making it a male-dominated field.
However, the careers in "Sales/Office", "Healthcare", and "Education" categories are female-dominated.

```{r}
#| label: earnings-job-category
#| fig-asp: 0.6
#| fig-width: 10

jobs_gender_plot <- jobs_gender |>
  mutate(
    major_category = fct_relevel(
      major_category,
      "Service",
      "Sales and Office",
      "Production, Transportation, and Material Moving",
      "Natural Resources, Construction, and Maintenance",
      "Management, Business, and Financial",
      "Healthcare Practitioners and Technical",
      "Education, Legal, Community Service, Arts, and Media",
      "Computer, Engineering, and Science"
      )
  )

ggplot(jobs_gender_plot, aes(x = major_category, fill = male_dom_workforce)) +
  geom_bar(position = "fill") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black") +
  labs(
    title = "Figure 2: 
Proportion of male-majority workforces across job categories",
    x = "Major job category",
    y = "Proportion of all observations",
    fill = "Workforce type"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Figure 3 displays the distribution of female wages as a percent of male wages across all workforces observed.
The unimodal, left-skewed distribution is concentrated below the dashed line, which indicates the ideal 1:1 ratio of female to male wages.
The median female wage as a percent of male wage is 85.2% with a standard deviation of 9.4%.

```{r}
#| label: histogram-percent-wage-male
#| fig-height: 4
#| fig-width: 6
jobs_gender |>
  ggplot(aes(x = wage_percent_of_male)) +
  geom_histogram(binwidth = 0.025) +
  geom_vline(xintercept = 1.0, linetype = "dashed", color = "black") +
  labs(
    title = "Figure 3: 
Distribution of female wages as a percent of male wages",
    x = "Female wages as a percent of male wages",
    y = "Count"
  ) +
  theme_minimal()
```

```{r}
#| label: median-sd-wage-gap
#| message: FALSE
#| echo: FALSE
eda_wage_gap <- jobs_gender |>
  summarize(
    median_wage_gap = median(wage_percent_of_male),
    sd_wage_gap = sd(wage_percent_of_male)
    )
```

We hypothesize that greater representation of females in the workforce will be correlated with higher female wages as a percent of male wages independent of and across all job categories.
More specifically, we predict the job categories with the highest proportion of observations that were male-majority workforces (see Figure 2), including the categories of "Computer, Engineering, and Science", "Natural Resources, Construction, and Maintenance", and "Production, Transportation, and Material Moving", will have lower female wages as a percent of male wages.

## Methodology

We chose to do an exploratory analysis first to better understand the relationships between variables.
Figure 1 displays the distributions of female wages as a percent of male wages across major job categories, suggesting the type of job could be a useful predictor of wage gaps due to the great variation in median and 25th to 75th quartiles.
Figure 2 highlighted our job categories of interest for our hypotheses: "Computer, Engineering, and Science", "Natural Resources, Construction, and Maintenance", and "Production, Transportation, and Material Moving".
Figure 3 displays our response variable of interest, female wages as a percent of male wages. We did not choose to use `year` as a variable because Table 1 did not indicate that it could be a promising predictor.

We chose to create three models to predict female wages as a percentage of male wages based on `major_category` and `percent_female`.
We chose to do linear models because our response variable is numerical and we have numerical and categorical predictor variables.
As we will discuss, Figure 4 demonstrates that a linear model is appropriate due to the moderately strong, positive, and linear relationship between our response variable and the percent of females in the workforce.
Our first model is a simple linear regression model that predicts the female wages as a percent of male wages based on the percent of the workforce that is female.
Our second model is an additive linear model which uses the predictors the percentage of the workforce that is female and the major job category.
Our third model is an interaction model between the percentage of the workforce that is female and the major job category.
Scatterplots visualize these relationships and the fitted lines.

We then selected the model with the highest $Adjusted~R^2$ and performed a hypothesis test on it using 1000 slope estimates. Hypothesis testing was appropriate to answer the question of whether greater representation of females in the workforce is correlated with higher female wages as a percent of male wages independent of and across all job categories, as there are two competing claims. The null hypothesis is that there is no relationship between representation in the workforce, job category, and wage gap, and the alternative hypothesis is that there is a relationship between these variables.

## Results

**Model 1**

According to Figure 4, the relationship between the percent of the workforce that is female and female wages as a percent of male wages appears to be moderately strong, positive, and linear.
Given this, we decided to fit a simple linear model with the percent of the female workforce as a predictor.

```{r}
#| label: scatterplot-simple-linear-model
#| fig-height: 4
#| fig-width: 6

ggplot(jobs_gender, aes(x = percent_female, y = wage_percent_of_male)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Figure 4: 
Wage inequality vs. gender representation in the workforce",
    x = "Percent of workforce that is female",
    y = "Female wages as percent of male wages",
  ) +
  theme_minimal()
```

```{r}
#| label: earnings-gap-workforce-gap-years-linear
#| message: FALSE
#| echo: false

simple_linear_fit <- linear_reg() |>
  fit(wage_percent_of_male ~ percent_female, data = jobs_gender)

```

$$\widehat{wage~percent~of~male} = 0.821 + 0.043 \times percent~female~workforce$$

Based on the linear model above, we predict that female employee wages as a percent of male wages when there are no female workers will be 82.1%, on average. In this context, the intercept is not meaningful because a gender wage gap cannot exist in a workforce with no female workers.
For every additional percentage point in percent of female employees in the workforce, we expect the female wage as a percent of male wage to be higher, on average, by 4.3 percentage points.

**Model 2**

Considering how the distribution of female wages as a percent of male wages (see Figure 1) varies based on the major job category, the following additive linear model uses both the percent of females in the workforce and the major job category as predictors.

```{r}
#| label: additive-linear-model-fit
#| message: FALSE
additive_linear_model <- linear_reg() |>
  fit(wage_percent_of_male ~ percent_female + major_category, data = jobs_gender)

```

```{r}
#| label: additive-scatterplot
#| fig-asp: 0.6
#| fig-width: 10

jobs_gender_additive_aug <- augment(additive_linear_model, new_data = jobs_gender)

ggplot(
  jobs_gender_additive_aug, 
  aes(x = percent_female, y = wage_percent_of_male, color = major_category)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = .pred), method = "lm") +
  labs(
    title = "Figure 5: 
Additive model of wage inequality vs. gender differences in workforce",
    subtitle = "by job category",
    x = "% of workforce that is female",
    y = "Female wages as % of male wages",
    color = "Major job category"
    ) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal()
```

$$
\small
\begin{aligned}
\widehat{wage~percent~of~male} &= 0.853 + 0.053 \times percent~female~workforce \\
&\quad -0.02 \times Education,~Legal,~Community~Service,~Arts,~and~Media \\
&\quad -0.026 \times Healthcare~Practitioners~and~Technical \\
&\quad -0.073 \times Management,~Business,~and~Financial \\
&\quad -0.004 \times Natural~Resources,~Construction,~and~Maintenance \\
&\quad -0.078 \times Production,~Transportation,~and~Material~Moving \\
&\quad -0.046 \times Sales~and~Office \\
&\quad -0.012 \times Service
\end{aligned}
$$

Based on the linear model above, we predict that female employee wages as a percent of male wages when there are no female workers and the job category is "Computer, Engineering, and Science" will be 85.3%, on average. In this context, the intercept is not meaningful because a gender wage gap cannot exist in a workforce with no female workers.
Holding all other variables constant, for every additional percentage point of female employees in the workforce, we expect the female wage as a percent of male wage to be higher, on average, by 5.3 percentage points.
The intercept adjustments for non-baseline major job categories are given above.
Holding all other variables constant, we expect an observation within the major job category of "Natural Resources, Construction, and Maintenance" to have a female wage as a percent of male wage that is lower, on average, by 0.4 percentage points.

**Model 3**

Considering the distribution of the percent of females in the workforce (see Figure 2) also varies based on the major job category, it is important to consider the interaction between these predictor variables.
Figure 6 depicts this interaction linear model.
The visual from this model displays the wage gap differences with respect to percentage of females in the workforce and their interaction with job categories. Each line has a different slope, with some going in the negative direction and others in the positive, indicating the direction of the relationship of that job category with the wage gap based on percentage of females in the workforce.

```{r}
#| label: scatterplot-interaction-lienar-model
#| fig-asp: 0.6
#| fig-width: 10

ggplot(jobs_gender, aes(
  x = percent_female, 
  y = wage_percent_of_male, 
  color = major_category)
  ) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Figure 6: 
Interaction model of wage inequality vs. gender differences in workforce",
    subtitle = "by job category",
    x = "% of workforce that is female",
    y = "Female wages as % of male wages",
    color = "Major job category"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal()
```

```{r}
#| label: interaction-linear-model-fit
#| message: FALSE
interaction_linear_model <- linear_reg() |>
  fit(wage_percent_of_male ~ percent_female * major_category, data = jobs_gender)

```

$$
\small
\begin{aligned}
\widehat{wage~percent~of~male} &= 0.909 -0.138 \times percent~female~workforce \\
&\quad -0.094 \times Education,~Legal,~Community~Service,~Arts,~and~Media \\
&\quad -0.214 \times Healthcare~Practitioners~and~Technical \\
&\quad -0.076 \times Management,~Business,~and~Financial \\
&\quad -0.048 \times Natural~Resources,~Construction,~and~Maintenance \\
&\quad -0.113 \times Production,~Transportation,~and~Material~Moving \\
&\quad -0.168 \times Sales~and~Office \\
&\quad -0.059 \times Service \\
&\quad +0.222 \times percent~female~workforce * Education,~Legal,~Community~Service,~Arts,~and~Media \\
&\quad +0.403 \times percent~female~workforce * Healthcare~Practitioners~and~Technical \\
&\quad +0.079 \times percent~female~workforce * Management,~Business,~and~Financial \\
&\quad +0.075 \times percent~female~workforce * Natural~Resources,~Construction,~and~Maintenance \\
&\quad +0.128 \times percent~female~workforce * Production,~Transportation,~and~Material~Moving \\
&\quad +0.302 \times percent~female~workforce * Sales~and~Office \\
&\quad +0.171 \times percent~female~workforce * Service \\
\end{aligned}
$$

Based on the linear model above, we predict that female employee wages as a percent of male wages when there are no female workers and the job category is "Computer, Engineering, and Science" will be 90.9%, on average. In this context, the intercept is not meaningful because a gender wage gap cannot exist in a workforce with no female workers.
Holding all other variables constant, for every additional percentage point of female employees in a "Computer, Engineering, and Science" category workforce, we expect the female wage as a percent of male wage to be lower, on average, by 13.8 percentage points. This finding is inconsistent with our prediction that workforces with higher female representation will have higher female wages as a percent of male wages.

The intercept and slope adjustments for non-baseline major job categories are given above. For our three male-majority job categories of interest, we have written the simplified equations below.

Computer, Engineering, and Science: 
$$\widehat{wage~percent~of~male} = 0.909 -0.138 \times percent~female~workforce$$
Natural Resources, Construction, and Maintenance: 
$$\widehat{wage~percent~of~male} = 0.861 -0.063 \times percent~female~workforce$$ 
Production, Transportation, and Material Moving: 
$$\widehat{wage~percent~of~male} = 0.796 -0.01 \times percent~female~workforce$$
Given these equations, our hypothesis that these male-majority specific job categories will have negative relationships with female wages as a percent of male wages is supported, regardless of the percent of females in the workforce.

**Selection of Linear Model**

```{r}
#| label: r2-model1
#| eval: false
#| message: FALSE
#glance(simple_linear_fit)
```

For the simple linear model, $Adjusted~R^2=0.012$ means that 1.2% of the variation in female wage as a percent of male wage is explained by the percent of females in the workforce.

```{r}
#| label: additive-linear-model-glance
#| message: FALSE
#| echo: FALSE
#glance(additive_linear_model)
```

For the additive linear moel, $Adjusted~R^2=0.097$ means that 9.7% of the variation in female wage as a percent of male wage is explained by the percent of females in the workforce and the major job category.

```{r}
#| label: interaction-linear-model-glance
#| echo: FALSE
#| eval: false
#glance(interaction_linear_model)
```

Finally, for the interaction linear model, $Adjusted~R^2=0.144$ means that 14.4% of the variation in female wage as a percent of male wage is explained by the interaction between the percent of females in the workforce and the major job category.

Because the $Adjusted~R^2$ of the interaction model is the highest, we have decided to select this model for further analysis.

**Hypothesis Test Using the Interaction Model**

Our null hypothesis, $H_0$, is that female wage as a percent of male wage is independent of both the percent of females in the workforce, the major job category, and the interaction between these predictor variables.
Our alternative hypothesis, $H_1$, is that there is a relationship between percent of females in the workforce, the major job category, and female wages as a percent of male wages.

We discuss the significance of the p-values of interest to our hypotheses in the Discussion section below.

```{r}
#| label: hypothesis-testing
#| message: FALSE
#| echo: FALSE
#| warning: FALSE

observed_fit <- jobs_gender |>
  specify(wage_percent_of_male ~ percent_female * major_category) |>
  fit()

set.seed(1234)

null_dist <- jobs_gender |>
  specify(wage_percent_of_male ~ percent_female * major_category) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()

null_dist_p_vals <- null_dist |>
  get_p_value(obs_stat = observed_fit, direction = "two-sided")
```
*Table 2*
```{r}
null_dist_p_vals
```

## Discussion

To reiterate, our research aims to explore how the gender pay gap varies with differences in percentage of females in the workforce and across major job categories. We compared three models’ adjusted $R^2$ values to determine their ability to predict female employee wages as a percent of male wages. All of the models were linear, as our response variable (`wage_percent_male`, female’s wages as a percent of male’s wages) is  a continuous response variable that would not be suited by a logistic model. Ultimately, we chose Model 3 to conduct our hypothesis testing, as it had the highest adjusted $R^2$, indicating it was the better fit for the model of the data while also accounting for the number of predictors in the model. 

From the p-values we obtained from our hypothesis testing and using a threshold value of 0.05, we reject the null hypothesis that female wage as a percent of male wage is independent of both the percent of females in the workforce, the major job category, and the interaction between these predictor variables. By rejecting the null, we are concluding that— while not necessarily a causal relationship– there *is* a relationship between percent of females in the workforce, the major job category, and female wages as a percent of male wages. This relationship varies by job category, and we found it interesting that almost all job categories demonstrated a p-value less than 0.05 (meaning we reject the null for all of these slope coefficients).

We can conclude that there is a statistically significant, non-zero  relationship between job category and female wage as a percent of male wage. We also can conclude that there is a statistically significant, non-zero relationship between the percent of females in a workforce and female wage as a percent of male wage. In regard to statistically significant conclusions for the interaction between job category and percent females in a workforce’s effects on female wage as a percent of male wage, the hypothesis test for five of the seven interaction terms showed a statistically significant result (p < 0.05). All of our majority-male job categories except for Natural Resources, Construction, and Management showed statistically significant p-values in both their interaction and direct coefficients (see Table 2). This implies a potential association between the interaction variables (job category and percent female of a workforce) and the outcome (female wage as a percent of male wage).

While we attempted to minimize limitations in our data and analysis, some were unavoidable. For one, the time period of data collection that we used in our analysis is 9 years old. As discussed in the introduction, pay gaps have not historically changed much over the years, but it could nonetheless be beneficial for the data used in our analysis to be more recent for the most accurate, up-to-date analysis possible. Another limitation of our data is the fact that only full-time jobs were reported; this leaves out a possibly interesting area of study including part-time job pay gap trends. Finally, as discussed in our introduction, some data points have “N/A” values for the `wage_percent_male` variable, which may have limited certain analyses– however, with the number of these being so low in comparison to the total number of observations, this limitation should not largely impact the outcomes of our analysis.
	
To improve our data , we think it could be beneficial to explore any trends in what data was missing or if missing data was clustered around certain time periods. If there were patterns in missing data (i.e., many `wage_percent_male` “N/A” values being in 2013), we could have adjusted our research question to more accurately suit the data we had available or calculated estimates if we had the numbers necessary. However, we do not expect this would make much of a difference in our outcomes, as “N/A” values were minimal in comparison to available data points.  
	
Some examples of potential issues related to the reliability and validity of our data include- but are not limited to-  recall bias and sampling error. With data being from the Census Bureau and the Bureau of Labor, responses were likely recorded using a survey. Surveys can lead to multiple issues with reliability: for one, individuals who participated may experience recall bias, mis-reporting their income or other factors of data. Further, not *all* working American people were included in this dataset, so while it may be generalizable, there is a chance for sampling error or misrepresentation of the wider population from this dataset. Overall, these impacts on reliability and validity are possible and unavoidable in the data we used for our analysis.

In the future, we think analyzing the effects of confounding factors such as race, age, and whether or not individuals were full- or part-time workers could be interesting. Gender may not be the only variable with discrepancies in pay, and it would be interesting to see how other demographic data impacts results. Further, we also think analyzing a more longitudinal dataset (i.e., from 2013 to present, 2025) could be intriguing in seeing how the gender pay gap has changed over time (if at all).


## References

Fry, R., & Aragão, C.
(2025, March 4).
Gender pay gap in U.S. has narrowed slightly over 2 decades.
Pew Research Center.
Retrieved April 21, 2025, from https://www.pewresearch.org/short-reads/2025/03/04/gender-pay-gap-in-us-has-narrowed-slightly-over-2-decades/#:\~:text=The%20U.S.%20Census%20Bureau%20has,Census%20Bureau%27s%20most%20recent%20analysis

The gender pay gap.
(n.d.).
American Association of University Women.
Retrieved April 21, 2025, from https://www.aauw.org/issues/equity/pay-gap/

Holder, M.
(2025, March 18).
Gender pay gap statistics 2025: A comprehensive analysis.
Equal Pay Today.
Retrieved April 21, 2025, from https://www.equalpaytoday.org/gender-pay-gap-statistics/

Kochhar, R.
(2023, March 1).
The enduring grip of the gender pay gap.
Pew Research Center.
Retrieved April 21, 2025, from https://www.pewresearch.org/social-trends/2023/03/01/the-enduring-grip-of-the-gender-pay-gap/

Occupational employment and wage statistics.
(n.d.).
U.S.
Bureau of Labor Statistics.
Retrieved April 21, 2025, from https://www.bls.gov/oes/oes_ques.htm

The simple truth about the gender pay gap.
(n.d.).
American Association of University Women.
Retrieved April 21, 2025, from https://www.aauw.org/resources/research/simple-truth/
