---
title: Differences in the Gender Pay Gap Across Male and Female-Dominated Workforces
subtitle: 2013-2016 trends in job category and female workforce representation
execute:
  echo: false
  message: false
  warning: false
---

Our research aims to analyze how women are marginalized in various job settings, specifically regarding how pay gaps fluctuate based on the type of job and how male-dominated a particular workforce is.

## Introduction and data

The gender pay gap has always been a relevant issue for women seeking equal workplace rights.
The disparities in wages are unfair, and accumulate to women only being able to save 70% as much as men on social security, on average (The Simple, n.d.).
Given the current political state and changes to DEI policies, this topic is crucial to understand and address from a statistical perspective.
In two decades, the gender pay gap has only narrowed by 4%, with women earning 81% as much as men in 2003 and 85% in 2024 (Fry & Aragão, 2025).
The data that we investigate in this study measures the pay gap for four years in between.
The gender pay gap is large and slowly moving, with estimations that it will take between 63 and 134 years to close (Holder 2025, The Gender, n.d.).
We are interested in how gender pay gaps vary across time from 2013 to 2016, varying percentages of female representation in the workforce, and specific major job categories, such as the Management, Business, and Financial sectors.

We hypothesize that major job categories that tend to be “male-dominated” will have a more significant male-female pay gap and fewer women working in that career.
Our research question asks, to what extent does the gender pay gap vary between male-dominated and female-dominated workforces and across major job categories?

We will primarily consider the variables of female wages as a percent of male wages (`wage_percent_of_male`), major job category (`major_category`), percentage of females in the workforce (`percent_female`), and year of data collection ranging from 2013 to 2016 (`year`).

All data was collected between 2013 and 2016 from statistics provided by the U.S.
Bureau of Labor through the U.S.
Census.
There are 2089 observations total, with 12 columns.
Each observation is a different combination of the year observed and occupation (i.e., “2013” and “Legislators” is one row/observation, and there are four rows of "Legislators" across all years).
301 occupations are shown throughout the 4 years, and 22 occupations are not represented in every year.
Considering that this data was scrubbed of any identifying factors aside from gender and was made public by the U.S.
Bureau of Labor Statistics, there are no ethical concerns with the data itself.

We transformed column names to provide more information and make data visualization and manipulation easier.
Those transformations include renaming the `total_earnings`, `total_earnings_male`, and `total_earnings_female` columns to `median_earnings`, `median_earnings_male`, and `median_earnings_female` given that the word "total" was misleading in communicating that the column represented the median earnings rather than all earnings for each group.
We also made percent_female and wage_percent_of_male into decimals rather than percentages.
Finally, to make the data easier to work with, we omitted observations with N/A values.

After our data cleaning, we have considered the following variables to be important in our data analysis: `year` (2013-2016), `major_category` (broad occupation category), `workers_male` and `workers_female` (estimated male or female full time workers above 16 years old), `percent_female` (percent of females for specific occupations), `total_earnings_male` which became `median_earnings_male` and `total_earnings_female` which became `median_earnings_female` (estimated male or female median earnings for full time workers over 16), and `wage_percent_of_male` (female wages as percent of male wages).

```{r}
#| label: load-packages
#| message: FALSE
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(dplyr)
jobs_gender <- read_csv("data/jobs_gender.csv")
```

```{r}
#| label: mutate-variables-to-decimal-form-and-rename
#| message: FALSE
# mutating percent_female to be displayed as a decimal
jobs_gender <- jobs_gender |>
  mutate(
# mutating percent_female to be displayed as a decimal
    percent_female = (percent_female * 0.01),
# mutating wage_percent_of_male to be displayed as a decimal
    wage_percent_of_male = (wage_percent_of_male * 0.01)
  ) |>
  rename(
# renaming total_earnings_male to median_earnings_male to be more clear
    median_earnings_male = total_earnings_male,
# renaming total_earnings_female to median_earnings_female to be more clear
    median_earnings_female = total_earnings_female,
# renaming total_earnings to median_earnings to be more clear
    median_earnings = total_earnings
  )
```

```{r}
#| label: creating-new-variables
#| message: FALSE
jobs_gender <- jobs_gender |>
# omit all observations with NA values.
  na.omit(jobs_gender) |>
  mutate(
# create male_dom_workforce as a binary factor where "Male dominated" means that there are more male workers than female workers, and "Female dominated" when there are more female workers.
    male_dom_workforce = as.factor(if_else(
      workers_male > workers_female, "Majority male", "Majority female")),
# create male_dom_earnings as a binary factor where Yes" means that male workers tend to be paid more than female workers, and "No" when female workers are usually paid more.
    male_dom_earnings = as.factor(if_else(
      median_earnings_male > median_earnings_female, "Yes", "No"))
    )
```

In order to begin exploring the data, we created the following three visualizations to understand which career categories show the most wage inequality, how gender representation varies by wage category, and how the variation in how much women are paid compared to men is shown.

The following plot visualizes the distribution of female wages as a percentage of male wages according to major job categories.

Based on the following visualization, we observed that all observations where women are paid equally or more than men are outside of the 75th percentile of observations, and are often considered outliers.
We also learned that computer, engineering, and science professions have the least versatility in wage inequality.

```{r}
#| label: scatterplot-wage-percentage-male
#| fig-asp: 0.5
#| fig-height: 8
jobs_gender |>
  ggplot(aes(x = wage_percent_of_male, color = major_category)) +
  geom_boxplot() +
  guides(color = guide_legend(reverse = TRUE)) +
  labs(
    title = "Figure 1: 
Distribution of female wages as percent of male wages",
    subtitle = "by major job category",
    x = "Female wages as % of male wages",
    color = "Major category"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(axis.ticks.y = element_blank(),
  axis.text.y = element_blank())
```

Male-dominated fields will be specified through the creation of two binary variables: `male_dom_workforce` labels observations with "Male majority" when the number of male workers is greater than the number of female workers and "Female majority" when there are more female workers and `male_dom_earnings` labels observations with "Yes" when the total male earnings are greater than the total female earnings and "No" when there are higher female earnings.
The following table demonstrates that, regardless of year, the amount of occupations showing male dominated earnings in this study remains around 300, and that occupations with female dominated earnings, while showing a slight increase over time, are still under 11.
Therefore, we do not put emphasis on year in this study as numbers remain fairly consistent.

```{r}
#| label: eda-male-dom-earnings
#| message: FALSE
eda_male_dom_earnings <- jobs_gender |>
  group_by(male_dom_earnings, year) |>
  count()
kableExtra::kable(eda_male_dom_earnings)
```

Figure 2 shows the proportion of occupations within the different major categories that reported having a male or female dominated work force.
The dashed line shows the 50% mark in the proportions, so any bar showing the blue male-majority line surpassing the dashed line has majority male-dominated work forces.
The graph shows that there is the least amount of female representation across observations in the job categories of natural resources, construction, and maintenance category making it a male dominated field.
However, the careers in sales/office, healthcare, and education are female-dominated.

```{r}
#| label: earnings-job-category
#| fig-asp: 0.6
#| fig-width: 10

jobs_gender_plot <- jobs_gender |>
  mutate(
    major_category = fct_relevel(
      major_category,
      "Service",
      "Sales and Office",
      "Production, Transportation, and Material Moving",
      "Natural Resources, Construction, and Maintenance",
      "Management, Business, and Financial",
      "Healthcare Practitioners and Technical",
      "Education, Legal, Community Service, Arts, and Media",
      "Computer, Engineering, and Science"
      )
  )

ggplot(jobs_gender_plot, aes(x = major_category, fill = male_dom_workforce)) +
  geom_bar(position = "fill") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black") +
  labs(
    title = "Figure 2: 
Male-dominated workforces",
    subtitle = "based on job categories",
    x = "Major job category",
    y = "Proportion of all observations",
    fill = "Workforce type"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Figure 3 displays the distribution of female wages as a percent of male wages across all workforces observed.
The unimodal, left-skewed distribution is concentrated below the dashed line, which indicates the ideal 1:1 ratio of female to male wages.
The median female wage as a percent of male wage is 85.2% with a standard deviation of 9.4%.

```{r}
#| label: histogram-percent-wage-male
#| fig-asp: 0.6
#| fig-width: 10
jobs_gender |>
  ggplot(aes(x = wage_percent_of_male)) +
  geom_histogram(binwidth = 0.025) +
  geom_vline(xintercept = 1.0, linetype = "dashed", color = "black") +
  labs(
    title = "Figure 3: 
Distribution of female wages as a percent of male wages",
    x = "Female wages as a percent of male wages",
    y = "Count"
  ) +
  theme_minimal()

eda_wage_gap <- jobs_gender |>
  summarize(
    median_wage_gap = median(wage_percent_of_male),
    sd_wage_gap = sd(wage_percent_of_male)
    )
kableExtra::kable(eda_wage_gap)
```

We hypothesize that greater representation of females in the workforce will be correlated with higher female wages as a percent of male wages independent of and across all job categories.
We also predict the job categories with the highest proportion of observations that were male-majority workforces (see Figure 2), including the categories of "Computer, Engineering, and Science", "Natural Resources, Construction, and Maintenance", and "Production, Transportation, and Material Moving", will have lower female wages as a percent of male wages.

## Methodology

We chose to do an exploratory analysis first to better understand the relationships between variables.
We created a bar chart that shows the proportions of workforces that do or do not have a median male wage higher than median female wages that is dodged based on the major job category.
EXPLAIN BRIEFLY WHY

We chose to create three models to predict female wages as a percentage of male wages based on `major_category` and `percent_female`.
We did not choose to use `year` as a variable because the scatterplot did not indicate that it could be a promising predictor.
We created a scatterplot that shows the female wages as a percent of male wages on the y-axis and the percentage of the workforce that is female on the x-axis with lines to designate the trends based on year (2013 vs. 2016) as we hypothesize that the percentage of the workforce that is female will be a major predictor of wage gaps.

We created three linear regression models.
Our first model is a simple linear regression model that predicts the female wages as a percent of male wages based on the percent of the workforce that is female.
Our second model is an additive linear model which uses the predictors the percentage of the workforce that is female and the major job category.
Our third model is an interaction model between the percentage of the workforce that is female and the major job category.
EXPLAIN WHY FOR EACH ----- Ellie

We then selected the model with the highest $Adjusted~R^2$ and performed hypothesis testing on it.
We thought hypothesis testing was appropriate to answer the question of whether greater representation of females in the workforce is correlated with higher female wages as a percent of male wages independent of and across all job categories as there are two competing claims: $H_0$ is that there is no relationship between the variables and $H_1$ is that there is a relationship between the variables.

SHOW HOW WE ARRIVED AT THE FINAL MODEL!!!

## Results

**Model 1**

According to Figure 4, the relationship between the percent of the workforce that is female and female wages as a percent of male wages appears to be moderately strong, positive, and linear.
Given this, we decided to fit a simple linear model with the percent of the female workforce as a predictor.

```{r}
#| label: scatterplot-simple-linear-model
#| fig-asp: 0.6
#| fig-width: 10

ggplot(jobs_gender, aes(x = percent_female, y = wage_percent_of_male)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Figure 4: 
Wage inequality vs. gender representation in the workforce",
    x = "% of workforce that is female",
    y = "Female wages as % of male wages",
  ) +
  theme_minimal()
```

```{r}
#| label: earnings-gap-workforce-gap-years-linear
#| message: FALSE

simple_linear_fit <- linear_reg() |>
  fit(wage_percent_of_male ~ percent_female, data = jobs_gender)

```

$$\widehat{wage~percent~of~male} = 0.821 + 0.043 \times percent~female~workforce$$

Based on the linear model above, we predict that female employee wages as a percent of male wages when there are no female workers will be 82.1%, on average.
For every additional percentage point in percent of female employees in the workforce, we expect the female wage as a percent of male wage to be higher, on average, by 4.3 percentage points.

```{r}
#| label: r2-model1
#| eval: false
#| message: FALSE
#glance(simple_linear_fit)
```

$Adjusted~R^2=0.012$ means that 1.2% of the variation in female wage as a percent of male wage is explained by the percent of females in the workforce.
$Adjusted~R^2$ was used in order to compare the effectiveness of the simple model with the following models with multiple predictors.

**Model 2**

Considering how the distribution of female wages as a percent of male wages (see Figure 1) varies based on the major job category, the following additive linear model uses both the percent of females in the workforce and the major job category as predictors.

```{r}
#| label: additive-linear-model-fit
#| message: FALSE
additive_linear_model <- linear_reg() |>
  fit(wage_percent_of_male ~ percent_female + major_category, data = jobs_gender)

```

```{r}
#| label: additive-scatterplot
#| fig-asp: 0.6
#| fig-width: 10

jobs_gender_additive_aug <- augment(additive_linear_model, new_data = jobs_gender)

ggplot(
  jobs_gender_additive_aug, 
  aes(x = percent_female, y = wage_percent_of_male, color = major_category)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = .pred), method = "lm") +
  labs(
    title = "Figure 5: 
Additive model of wage inequality vs. gender differences in workforce",
    subtitle = "by job category",
    x = "% of workforce that is female",
    y = "Female wages as % of male wages",
    color = "Major job category"
    ) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal()
```

$$
\small
\begin{aligned}
\widehat{wage~percent~of~male} &= 0.853 + 0.053 \times percent~female~workforce \\
&\quad -0.02 \times Education,~Legal,~Community~Service,~Arts,~and~Media \\
&\quad -0.026 \times Healthcare~Practitioners~and~Technical \\
&\quad -0.073 \times Management,~Business,~and~Financial \\
&\quad -0.004 \times Natural~Resources,~Construction,~and~Maintenance \\
&\quad -0.078 \times Production,~Transportation,~and~Material~Moving \\
&\quad -0.046 \times Sales~and~Office \\
&\quad -0.012 \times Service
\end{aligned}
$$

Based on the linear model above, we predict that female employee wages as a percent of male wages when there are no female workers and the job category is Computer, Engineering, and Science will be 85.3%, on average.
Holding all other variables constant, for every additional percentage point of female employees in the workforce, we expect the female wage as a percent of male wage to be higher, on average, by 5.3 percentage points.
The intercept adjustments for non-baseline major job categories are given above.
Holding all other variables constant, we expect an observation within the major job category of Education, Legal, Community Service, Arts, and Media to have a female wage as a percent of male wage that is lower, on average, by 2.0 percentage points.

```{r}
#| label: additive-linear-model-glance
#| message: FALSE
#| echo: FALSE
#glance(additive_linear_model)
```

$Adjusted~R^2=0.097$ means that 9.7% of the variation in female wage as a percent of male wage is explained by the percent of females in the workforce and the major job category.

**Model 3**

Considering the distribution of the percent of females in the workforce (see Figure 2) also varies based on the major job category, it is important to consider the interaction between these predictor variables.
Figure 6 depicts this interaction linear model.

```{r}
#| label: scatterplot-interaction-lienar-model
#| fig-asp: 0.6
#| fig-width: 10

ggplot(jobs_gender, aes(
  x = percent_female, 
  y = wage_percent_of_male, 
  color = major_category)
  ) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Figure 6: 
Interaction model of wage inequality vs. gender differences in workforce",
    subtitle = "by job category",
    x = "% of workforce that is female",
    y = "Female wages as % of male wages",
    color = "Major job category"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal()
```

```{r}
#| label: interaction-linear-model-fit
#| message: FALSE
interaction_linear_model <- linear_reg() |>
  fit(wage_percent_of_male ~ percent_female * major_category, data = jobs_gender)

```

$$
\small
\begin{aligned}
\widehat{wage~percent~of~male} &= 0.909 -0.138 \times percent~female~workforce \\
&\quad -0.094 \times Education,~Legal,~Community~Service,~Arts,~and~Media \\
&\quad -0.214 \times Healthcare~Practitioners~and~Technical \\
&\quad -0.076 \times Management,~Business,~and~Financial \\
&\quad -0.048 \times Natural~Resources,~Construction,~and~Maintenance \\
&\quad -0.113 \times Production,~Transportation,~and~Material~Moving \\
&\quad -0.168 \times Sales~and~Office \\
&\quad -0.059 \times Service \\
&\quad +0.222 \times percent~female~workforce * Education,~Legal,~Community~Service,~Arts,~and~Media \\
&\quad +0.403 \times percent~female~workforce * Healthcare~Practitioners~and~Technical \\
&\quad +0.079 \times percent~female~workforce * Management,~Business,~and~Financial \\
&\quad +0.075 \times percent~female~workforce * Natural~Resources,~Construction,~and~Maintenance \\
&\quad +0.128 \times percent~female~workforce * Production,~Transportation,~and~Material~Moving \\
&\quad +0.302 \times percent~female~workforce * Sales~and~Office \\
&\quad +0.171 \times percent~female~workforce * Service \\
\end{aligned}
$$

Based on the linear model above, we predict that female employee wages as a percent of male wages when there are no female workers and the job category is Computer, Engineering, and Science will be 90.9%, on average.
This makes sense given the well-known disparities in women's wages in STEM fields.
Holding all other variables constant, for every additional percentage point of female employees in a Computer, Engineering, and Science category workforce, we expect the female wage as a percent of male wage to be lower, on average, by 13.8 percentage points.
This intercept is inconsistent with our prediction that workforces with higher female representation will have higher female wages as a percent of male wages.
The intercept and slope adjustments for non-baseline major job categories are given above.
For our three male-majority job categories of interested, we have written the simplified equations below.
Computer, Engineering, and Science: $$\widehat{wage~percent~of~male} = 0.909 -0.138 \times percent~female~workforce$$ Natural Resources, Construction, and Maintenance: $$\widehat{wage~percent~of~male} = 0.861 -0.063 \times percent~female~workforce$$ Production, Transportation, and Material Moving: $$\widehat{wage~percent~of~male} = 0.796 -0.01 \times percent~female~workforce$$

```{r}
#| label: interaction-linear-model-glance
#| message: FALSE
#| eval: false
#glance(interaction_linear_model)
```

$Adjusted~R^2=0.144$ means that 14.4% of the variation in female wage as a percent of male wage is explained by the interaction between the percent of females in the workforce and the major job category.

**Selection of Linear Model**

Because the $Adjusted~R^2$ of the interaction model is the highest, we have decided to select this model for further analysis.

**Hypothesis Test Using the Interaction Model**

Our null hypothesis is that female wage as a percent of male wage is independent to both the percent of females in the workforce, the major job category, and the interaction between these predictor variables.
Our alternative hypothesis is that there is a relationship between percent of females in the workforce, the major job category, and female wages as a percent of male wages.

The p-value is 0, which means that we should reject the null hypothesis as assuming the null hypothesis is true, the probability of observing what we did is 0.

```{r}
#| label: hypothesis-testing
#| message: FALSE
#| echo: FALSE
#| warning: FALSE

observed_fit <- jobs_gender |>
  specify(wage_percent_of_male ~ percent_female * major_category) |>
  fit()

set.seed(1234)

null_dist <- jobs_gender |>
  specify(wage_percent_of_male ~ percent_female * major_category) |>
  hypothesize(null = "independence") |>
  generate(reps = 100, type = "permute") |>
  fit()

null_dist_p_vals <- null_dist |>
  get_p_value(obs_stat = observed_fit, direction = "two-sided")
```

## Discussion

In this section you’ll include a summary of what you have learned about your research question along with statistical arguments supporting your conclusions.
In addition, discuss the limitations of your analysis and provide suggestions on ways the analysis could be improved.
Any potential issues pertaining to the reliability and validity of your data and appropriateness of the statistical analysis should also be discussed here.
Lastly, this section will include ideas for future work.

## References

Fry, R., & Aragão, C.
(2025, March 4).
Gender pay gap in U.S. has narrowed slightly over 2 decades.
Pew Research Center.
Retrieved April 21, 2025, from https://www.pewresearch.org/short-reads/2025/03/04/gender-pay-gap-in-us-has-narrowed-slightly-over-2-decades/#:\~:text=The%20U.S.%20Census%20Bureau%20has,Census%20Bureau%27s%20most%20recent%20analysis

The gender pay gap.
(n.d.).
American Association of University Women.
Retrieved April 21, 2025, from https://www.aauw.org/issues/equity/pay-gap/

Holder, M.
(2025, March 18).
Gender pay gap statistics 2025: A comprehensive analysis.
Equal Pay Today.
Retrieved April 21, 2025, from https://www.equalpaytoday.org/gender-pay-gap-statistics/

Kochhar, R.
(2023, March 1).
The enduring grip of the gender pay gap.
Pew Research Center.
Retrieved April 21, 2025, from https://www.pewresearch.org/social-trends/2023/03/01/the-enduring-grip-of-the-gender-pay-gap/

Occupational employment and wage statistics.
(n.d.).
U.S.
Bureau of Labor Statistics.
Retrieved April 21, 2025, from https://www.bls.gov/oes/oes_ques.htm

The simple truth about the gender pay gap.
(n.d.).
American Association of University Women.
Retrieved April 21, 2025, from https://www.aauw.org/resources/research/simple-truth/
